# Cartogram

Following:

* [https://r-graph-gallery.com/331-basic-cartogram.html](https://r-graph-gallery.com/331-basic-cartogram.html) 
* [https://cran.r-project.org/web/packages/cartogram/readme/README.html](https://cran.r-project.org/web/packages/cartogram/readme/README.html)

## Call Libraries

```{r}
#| output: false
#| code-fold: true

library(rnaturalearth) # natural earth data

library(ggplot2) # beautiful maps

library(dplyr) # data wrangling

library(sf) # simple (spatial) features

library(cartogram) # cartograms!

```

## Remove Scientific Notation

```{r}
#| code-fold: true

options(scipen = 999) # high 'penalty' for scientific notation

```

## Get Map Data From `rnaturalearth`

```{r}
#| code-fold: true

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf")  # as sf object
  
```

## Make A Basic Map

We make a basic map, reading it into an object called `mymap.` We then *replay* `mymap.`

```{r}
#| code-fold: true

mymap <- ggplot(mapdata) + # the data I am mapping
  geom_sf() # the geometry I am using

mymap # replay my map

```

## Make And Plot The Cartogram

### Make The Cartogram Data

```{r}
#| code-fold: true

mapdata_merc <- st_transform(mapdata, 
                             3857) # transform to Mercator (why?)

# make a cartogram
# each iteration takes a LONG time
# fewer iterations help the time
# but each iteration contributes to the 'distortion'
# and makes a more 'cartogram'-like cartogram

start_time <- Sys.time() # time this step

mapdata_cartogram <- cartogram_cont(mapdata_merc, 
                                  "pop_est", 
                                  itermax = 7)

end_time <- Sys.time()

end_time - start_time

```

### Basic Cartogram

```{r}
#| code-fold: true

ggplot(mapdata_cartogram) + 
  geom_sf()

```

::: {.callout-tip}
### Why Does Antarctica Look So Strange? How To Fix This?

The key is to run this `dplyr` code to remove Antarctica.

```{r}
#| eval: true
#| code-fold: true

mapdata_cartogram <- mapdata_cartogram %>% 
  dplyr::filter(! continent == "Antarctica")

```
:::

### More Advanced Cartogram

```{r}
#| code-fold: true

ggplot(mapdata_cartogram) + 
  geom_sf(aes(fill = pop_est)) + # fill is population estimate
  scale_fill_viridis_c(name = "population",
                       option = "turbo") + # beautiful colors
  theme_void()

```


