# Geocoding {#sec-geocoding}

## Call Libraries

```{r}
#| message: false
#| code-fold: true

library(tidygeocoder) # geocoding

library(dplyr) # for %>% operator

library(readr) # import CSV

library(DT) # nice tables

```

## Get Data To Be Geocoded

```{r}
#| code-fold: true
#| output: false

simulated_address_data <- read_csv("simulated-address-data/simulated-address-data.csv")

DT::datatable(simulated_address_data,
              extensions = 'Buttons', 
              options = list(
                dom = 'Bfrtip',
                buttons = c('copy', 
                            'csv', 
                            'print'))) # nice table

```

## Concatenate Addresses

```{r}
#| code-fold: true

simulated_address_data$address <- paste(simulated_address_data$street,
                                        ", ",
                                        simulated_address_data$city,
                                        ", ",
                                        simulated_address_data$state)

DT::datatable(simulated_address_data,
              extensions = 'Buttons', 
              options = list(
                dom = 'Bfrtip',
                buttons = c('copy', 
                            'csv', 
                            'print'))) # nice table

```

## Geocode

> ArcGIS geocoding has LOW success rate with this data

> You will want to find a process with HIGH success rate

> You could also try batchgeo -> KML -> Latitude/Longitude

```{r}
#| code-fold: true

geocoded_data <- simulated_address_data %>% 
  tidygeocoder::geocode(address, 
                        method = 'arcgis', 
                        lat = latitude, 
                        long = longitude)

DT::datatable(geocoded_data,
              extensions = 'Buttons', 
              options = list(
                dom = 'Bfrtip',
                buttons = c('copy', 
                            'csv', 
                            'print'))) # nice table

```

::: {.callout-important title="Geocoding Can Make Errors!"}
NB that for whatever reason, the geocoder has made a mistake: Agency Z has been placed in the Southern Hemisphere. Geocoding can be an error prone process and requires careful inspection of your tabular and mapped data.

A geocoder may also be unable to geocode some of your data. Low success rates are not uncommon, and you may have to work hard to ensure that the majority, or all, of your data are geocoded.
:::

> Geocoded data can then be mapped using procedures outlined in @sec-location-data.



