# Introduction to Mapping and GIS

```{r}
#| output: false
#| echo: false

library(rnaturalearth) # natural earth data

library(sf) # simple (spatial) features

library(sp) # spatial data

library(leaflet) # beautiful maps

library(plotly) # interactive graphics

```

```{r}
#| label: fig-world
#| fig-cap: "Countries of the World"
#| fig-height: 3
#| echo: false
#| column: body

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf") # as sf object

mapdata <- mapdata %>% 
  sf::st_transform('+proj=longlat +datum=WGS84') # re-project

leaflet() %>%
  setView(0, 0, zoom = 1.5) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addPolygons(data = mapdata, 
              fillOpacity = .0, 
              weight = 2,
              color = "#009edb", 
              label = mapdata$name,
              highlightOptions = highlightOptions(color = "red", 
                                                  weight = 2,
                                                  bringToFront = TRUE)) 

```

Mapping and GIS have become an increasingly important part of social research, advocacy, and program and policy development. This website contains materials for an *in development* course on *Mapping and GIS*.

The content of this site is built around the use of R software, though many of the concepts are applicable across GIS and mapping applications such as R, Stata, ArcGIS and QGIS.










