---
title: "Introduction"
---

```{r}
#| output: false
#| echo: false

library(sf) # simple (spatial) features

library(leaflet) # beautiful maps

library(plotly) # interactive graphics

```

```{r}
#| label: fig-world
#| fig-cap: "Countries of the World"
#| fig-height: 3
#| echo: false

global_data <- read_sf("./shapefiles/wrld_simpl/wrld_simpl.shp")

  leaflet() %>%
    setView(0, 0, zoom = 1.5) %>%
    addProviderTiles(providers$Esri.WorldImagery) %>%
    addPolygons(data = global_data, 
                fillOpacity = .0, 
                weight = 2,
                color = "#009edb", 
                label = global_data$NAME,
                highlightOptions = highlightOptions(color = "red", 
                                                    weight = 2,
                                                    bringToFront = TRUE)) 

```

Mapping and GIS have become an increasingly important part of social research, advocacy, and program and policy development. This website contains materials for an *in development* course on *Mapping and GIS*.

Much of the content of this site is built around the use of R software, though many of the concepts are applicable across GIS and mapping applications such as R, Stata, ArcGIS and QGIS.










