---
title: "Latitude and Longitude"
---

# Introduction

Latitude and longitude are coordinates for locating objects on earth.

* Latitude represents distance from the *equator*. 0 latitude is at the equator. +90 latitude, or 90N, is at the North Pole. -90 latitude, or 90S, is at the South Pole.
* Longitude represents distance from the *prime meridian*. 0 longitude is at the prime meridian. +180 and -180, or 180E and 180W, meet at the other side of the world.

# Call the Libraries

```{r}
#| message: false
#| code-fold: true

library(plotly)

```

# Generate Some Random Coordinates

```{r}
#| code-fold: true

set.seed(3846) # random seed

N <- 10 # number of points

# latitude from -90 to +90

latitude <- runif(N, min = -90, max = 90) 

# longitude from + -180 to + 180

longitude <- runif(N, min = -180, max = 180) 

# 1st point reset to 0, 0

latitude[1] <- 0 # equator

longitude[1] <- 0 # prime meridian

latitude[2] <- 42 # Ann Arbor-ish

longitude[2] <- -83.5 # Ann Arbor-ish

# label

label <- LETTERS[1:N] # label with letters of alphabet

# dataframe

mydata <- data.frame(latitude, longitude, label)

mydata # replay

```

# Map The Coordinates

```{r}
#| code-fold: true

g <- list(lonaxis = list(showgrid = T, # geographic parameters
                         gridcolor = "lightblue"), 
          lataxis = list(showgrid = T, 
                         gridcolor = "lightblue"),
          showland = TRUE,
          landcolor = toRGB("lightgrey"))

mymap <- plot_geo(mydata) %>%
  add_markers(x = ~longitude,
              y = ~latitude,
              color = ~label,
              colors = "Spectral",
              marker = list(size = 15)) %>%
  layout(title = "Randomly Generated Coordinates",
         geo = g)

mymap # replay

```

