# Map Projections {#sec-projections}

## Introduction

Map projections exist because we are trying to take the round globe of the earth, and project it onto a 2 dimensional surface. Because a spherical globe can not be projected onto a flat surface without *some* distortion, different projections make different choices about the kind of distortion involved.

```{r}
#| echo: false
#| message: false
#| fig-cap: 
#|   - "A Flat Map"
#|   - "A Globe"
#| label: globe-and-flat-map
#| layout-ncol: 2

library(rnaturalearth) # natural earth data

library(ggplot2) # beautiful maps

library(dplyr) # data wrangling

library(sf) # simple (spatial) features

# library(patchwork) # stitch ggplots together

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf") # as sf object

mymap <- ggplot(mapdata) + # the data I am mapping
  geom_sf() + # the geometry I am using
  theme_minimal() +
  theme(axis.text.x = element_blank())

mymap # + coord_sf(crs="+proj=robin") 

mymap + coord_sf(crs="+proj=ortho")
  
```

## Get The Data

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf") # as sf object

```

## A Basic Map

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mymap <- ggplot(mapdata) + # the data I am mapping
  geom_sf() + # the geometry I am using
  theme_minimal() + # minimal theme
  theme(axis.text.x = element_blank()) # no longitude labels

mymap # replay

```

## Map Projections

### Globe (Orthographic)

::: {.callout-tip}
## Key Idea

An *orthographic* projection represents the globe as a 3 dimensional view.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mymap + coord_sf(crs="+proj=ortho")

```

### Mercator 

::: {.callout-tip}
## Key Idea

A *Mercator* projection represents the earth with perpendicular latitude and longitude. This projection can be helpful in some kinds of navigation, but areas of landmasses are distorted. As one approaches the poles, landmasses are over-sized, while landmasses closer to the equator are under-emphasized. Thus, this projection is often seen as one that does not properly acknowledge the size of countries in the Global South.

Antarctica cannot be correctly mapped with the Mercator projection. Thus, we need to employ a slightly more complicated procedure, removing Antarctica from the data before plotting.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mercator_data <- mapdata %>% 
  filter(name != "Antarctica") # remove Antarctica

ggplot(mercator_data) + # the data I am mapping
  geom_sf() + # the geometry I am using
  coord_sf(crs = 3857) + # Mercator
  theme_minimal() + # minimal theme
  theme(axis.text.x = element_blank()) # no longitude labels

```

### Mollweide

::: {.callout-tip}
## Key Idea

The *Mollweide* projection is an *equal area* projection. As a consequence, latitude and longitude lines are not perpendicular, and the shapes of some landmasses may appear to be distorted.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mymap + coord_sf(crs="+proj=moll")

```

### Robinson

::: {.callout-tip}
## Key Idea

The *Robinson* projection is an attempt to compromise between equal areas and a natural looking map.
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mymap + coord_sf(crs="+proj=robin") 

```







