# Using Data From `rnaturalearth` {#sec-rnaturalearth}

`rnaturalearth` is a source library for data various types of global mapping data.

## Call Libraries

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false
#| warning: false

library(rnaturalearth) # natural earth data

library(ggplot2) # beautiful maps

library(dplyr) # data wrangling

library(sf) # simple (spatial) features

```

## Get `mapdata` As `sf`

`ne_countries` stands for *Natural Earth Countries*. 

I use `ne_countries()` to create the `mapdata` dataset as an `sf` object (@sec-sf).

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

mapdata <- ne_countries(scale = "medium", # medium scale
                        returnclass = "sf") # as sf object

```

## Map

### Simple Basic Map

I make a simple map of this `sf` object with `ggplot` (@sec-ggplot-map).

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

ggplot(mapdata) + # the data I am mapping
  geom_sf() # the geometry I am using

```

### More Complicated Map


```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| message: false

ggplot(mapdata) + # the sf data that I am mapping
  geom_sf(aes(fill = income_grp)) + # what goes on the map: FILL
  scale_fill_viridis_d(name = "Income Group", # beautiful colors
                       option = "viridis") +
  labs(title = "Countries of the World") + # labels
  theme_minimal() # minimal theme

```










