# Shapefiles {#sec-shapefiles}

## Introduction 

Shapefiles are a spatial data format originally developed by [ESRI](https://www.esri.com/en-us/home).

Shapefiles come in three major types:

* **points**, to represent point features such as individual or agency locations. At larger scales, cities or towns might also be points.
* **lines**️, to represent line features such as roads, trails, or rivers.
* **polygons**, to represent polygon features such as outlines of cities, states, or countries.

## Shape*files* Are Actually *Collections* of Files

Shape*files* are actually a *set* or *collection* of associated files, all with the same name, and all in the same directory, but different suffixes. 

R--and many other software programs--generally reference the `*.shp` file of the shapefile.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

list.files("./shapefiles/a2trees")

```

## Call Libraries

```{r}
#| message: false
#| code-fold: true
#| code-summary: "Show the code"

library(ggplot2) # beautiful graphs

library(sf) # simple (spatial) features

```

## Open Shapefiles

```{r, eval=TRUE, warning=FALSE}
#| code-fold: true
#| code-summary: "Show the code"

city_boundary <- read_sf("./shapefiles/AA_City_Boundary/AA_City_Boundary.shp")

buildings <- read_sf("./shapefiles/AA_Building_Footprints/AA_Building_Footprints.shp")

trees <- read_sf("./shapefiles/a2trees/AA_Trees.shp")

parks <- read_sf("./shapefiles/AA_Parks/AA_Parks.shp")

university <- read_sf("./shapefiles/AA_University/AA_University.shp")

WashtenawRoads <- read_sf("./shapefiles/Roads/RoadCenterlines.shp")

AnnArborRoads <- st_crop(WashtenawRoads, 
                         city_boundary) # crop to only get A2 roads

# watersheds <- read_sf("./shapefiles/watersheds/Watersheds.shp")

```

## Use `ggplot` to Map the Shapefiles

```{r}
#| code-fold: true
#| code-summary: "Show the code"

ggplot(city_boundary) + # initial sf data
  # geom_sf(data = buildings,
  #         fill = "lightgrey") +
  geom_sf(data = AnnArborRoads, # first layer: Ann Arbor roads 
          color = "lightgrey") +
  geom_sf(color = "red", alpha = .5) + # second layer: city boundary
  geom_sf(data = trees, # third layer: trees
          size = .01,
          color = "darkgreen") +
  labs(title = "Demonstration of Shapefiles",
       subtitle = "Green Trees Are Points \nGrey Roads are Lines \nRed City Boundary is Polygon") +
  theme_minimal() +
  theme(axis.text = element_text(size = rel(.5)))

```

