# Using Data From `WDI` {#sec-WDI}

## Background

The [World Bank](http://www.worldbank.org/) collects statistical information from countries around the world.  A particularly useful data set is the **W**orld **D**evelopment **I**ndicators [(WDI)](http://data.worldbank.org/data-catalog/world-development-indicators) which are country level statistical information from around the world.

Using `library(WDI)` you can download indicator data directly from the World Bank and read it into a data set.

## Call Libraries

```{r}
#| message: false
#| warning: false
#| code-fold: true

library(WDI) # for accessing World Bank data

library(dplyr) # data wrangling

```

## Get Some Data From the World Development Indicators (WDI)

```{r}
#| eval: false
#| code-fold: true

# get names of specific indicators from WDI Data Catalog

WorldBankData <- WDI(country="all", 
                     indicator=c("SI.POV.GINI", # Gini
                                 "NY.GDP.PCAP.CD", # GDP
                                 "SE.ADT.LITR.ZS", # adult literacy
                                 "SP.DYN.LE00.IN", # life expectancy
                                 "SP.POP.TOTL", # population
                                 "SN.ITK.DEFC.ZS"), # undernourishment
                     start = 2023, 
                     end = 2023, 
                     extra = TRUE) 


save(WorldBankData, file="WorldBankData.RData")

```

# Rename Some Variables

```{r}
#| eval: false
#| code-fold: true

# think about renaming some variables with more intuitive names
# e.g....

# rename some variables with dplyr (just copy and paste your indicators)

WorldBankData <- dplyr::rename(WorldBankData, 
                        GDP = NY.GDP.PCAP.CD,
                        adult_literacy = SE.ADT.LITR.ZS,
                        life_expectancy = SP.DYN.LE00.IN, 
                        population = SP.POP.TOTL,
                        Gini = SI.POV.GINI,
                        undernourishment = SN.ITK.DEFC.ZS)

save(WorldBankData, file="WorldBankData.RData")

```

## Look At The Data

```{r}
#| code-fold: true

load("WorldBankData.RData") # load the data

head(WorldBankData)

```


