---
title: "Latitude and Longitude"
subtitle: "An evolving tutorial."
author: "Andy Grogan-Kaylor"
date: "today"
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Show the code"
    theme:
      light: united
      dark: vapor
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{css, echo=FALSE}

blockquote {
  color: black;
  border-left: 2px solid #FFCB05; 
  padding: 0.5em 10px;
}
  
```

# Introduction

Latitude and longitude are coordinates for locating objects on earth.

* Latitude represents distance from the *equator*. 0 latitude is at the equator. +90 latitude, or 90N, is at the North Pole. -90 latitude, or 90S, is at the South Pole.
* Longitude represents distance from the *prime meridian*. 0 longitude is at the prime meridian. +180 and -180, or 180E and 180W meet at the other side of the world.

# Call the Libraries

```{r, message=FALSE}

library(plotly)

```

# Generate Some Random Coordinates

```{r}

N <- 10 # number of points

# latitude from -90 to +90

latitude <- runif(N, min = -90, max = 90) 

# longitude from + -180 to + 180

longitude <- runif(N, min = -180, max = 180) 

# 1st point reset to 0, 0

latitude[1] <- 0

longitude[1] <- 0

# label

label <- LETTERS[1:N] # label with letters of alphabet

# dataframe

mydata <- data.frame(latitude, longitude, label)

mydata # replay

```

# Map The Coordinates

```{r}

g <- list(lonaxis = list(showgrid = T, # geographic parameters
                         gridcolor = "lightblue"), 
          lataxis = list(showgrid = T, 
                         gridcolor = "lightblue"),
          showland = TRUE,
          landcolor = toRGB("lightgrey"))

mymap <- plot_geo(mydata) %>%
  add_markers(x = ~longitude,
              y = ~latitude,
              color = ~label,
              colors = "Spectral",
                  marker = list(size = 15)) %>%
  layout(title = "Randomly Generated Coordinates",
         geo = g)

mymap # replay

```

